Covid-19 infection and attributable mortality in residents of UK care homes: Cohort study using electronic records (March-June 2020)

### 2020-07-08 manuscript/manuscript.Rmd
manuscript v0.5

---
title: "Incidence of confirmed Covid-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)

@@

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-14")
date_min_tallies = as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-14")

@@

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(covid_first_confirmed, na.rm = T), source="DAT")
) %>% 
  dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(first_case_date)) %>% 
  tidyr::crossing(tibble(outbreak = c(0, 1)))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  )) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

@@

tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

supp_epi_DAT <- table_nums(
  name = "supp_epi_DAT",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections amongst residents according to Datix incident reports (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))

@@

tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = paste0("Number and proportion of homes without confirmed cases by UK region (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

@@

fig_KM_CT_residents <- fig_nums(
  name = "fig_KM_CT_residents",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in residents according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

fig_KM_staff <- fig_nums(
  name = 'fig_KM_staff',
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infection in care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")"))

@@

**Findings:** `r formatbm(desc_residents$CT_total_symptomatic)` residents developed Covid-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had confirmed infections. The incidence of confirmed infection in residents and staff respectively was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days across all care homes. This includes `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) which had not reported a single confirmed infection or Covid-related death.

Among `r formatbm(desc_residents$DAT_total_residents)` private residents, `r formatbm(desc_residents$DAT_total_deaths)` all-cause deaths occurred either in homes or in hospital (`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%]), including `r desc_residents$DAT_CFR$y` deaths of residents who had had a confirmed infection. This suggests an all-cause case-fatality rate of `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%], of which an estimated `r tbrounding(((desc_residents$DAT_CFR_RR - 1)/desc_residents$DAT_CFR_RR)*100)` [`r tbrounding(((desc_residents$DAT_CFR_RR_lower - 1)/desc_residents$DAT_CFR_RR_lower)*100)`; `r tbrounding(((desc_residents$DAT_CFR_RR_upper - 1)/desc_residents$DAT_CFR_RR_upper)*100)`] is attributable to confirmed SARS-CoV-2 infections. Independently, care home reports of Covid-related deaths suggest that `r desc_residents$CT_total_cdeath`/`r formatbm(desc_residents$CT_total_residents)` residents (`r tbrounding(desc_residents$CT_p_cdeath)`% [`r tbrounding(desc_residents$CT_p_cdeath_lower)`; `r tbrounding(desc_residents$CT_p_cdeath_upper)`]) died of a cause directly attributable to Covid-19. In addition to male gender and older age, confirmed asymptomatic (HR = 3·3 [1·9; 5·6]) and confirmed symptomatic (HR = 12 [10; 15]) infection were independently associated with all-cause mortality, by comparison with asymptomatic residents living in homes with no record of confirmed infections. All-cause mortality for uninfected residents was 93% higher in care homes with outbreaks compared to those without outbreaks (adjusted HR = 1·9 [1·7; 2·2]).

@@

Care home residents are particularly vulnerable to Covid-19 infection due to their advanced age and high prevalence of comorbidity,[@Zhou2020] and this is compounded by frequent exposure to infection through close contact with infected staff members, other residents and contaminated surfaces in the care home setting. In the UK, an estimated 420,000 residents lived in approximately 11,000 homes for the elderly.[@cqcdata2020] At the peak of the pandemic the number of deaths in care home residents increased threefold over the equivalent period in 2019.[@onsDR2020]

However, this is likely to underestimate of the true number of deaths in care home residents that were caused by the pandemic, since it only includes deaths in which Covid-19 was directly implicated. Death certifications in England and Wales between 28th December 2019 and 12th June 2020 suggest that two thirds of the 29,393 excess deaths recorded in care home residents (calculated relative to equivalent period in the previous year) involved Covid-19.[@onsDR2020] As part of efforts to mitigate the anticipated second wave of infection, it is critical to understand the ratio of direct to indirect deaths in care home residents. This will inform decisions around instigating lockdown measures to protect residents and staff from infection, whilst recognising the potential unintended consequences of interrupting usual care.

Disentangling direct and indirect deaths from Covid-19 requires accurate information on the incidence and prevalence of infection among staff and residents linked to individual-level clinical outcomes. There is no syndromic surveillance for infection in care homes in England, and universal testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established in care homes in England until 11th May 2020, when the government launched the whole care home testing programme.[@dhscTesting2020] Prior to this, testing was only available for care home residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations and within a maximum of five tests per care home. Consequently national estimates of incidence and prevalence based on test results obtained prior to 11th May will substantially underestimate the burden of infection in care home residents and staff.

Outbreak investigations are useful to estimate the proportion of residents and staff who become infected during an outbreak,[@McMichael2020; @Dora2020] but they provide limited insight into the relationship between infection and mortality across the entire pandemic period, and cannot be generalised to care homes that have not reported any outbreaks. A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long-term care facilities. Mortality rates in residents with confirmed Covid-19 infection ranged from 0 to 10%, with case fatality rates between 0 and 34%, however this review did not include estimates indirect mortality.

@@

## Aggregate outbreak surveillance counts (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")`)

@@

## Individual-level residents records (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")`)

@@

Incidence proportions and rates were calculated using the aggregate outbreak counts, as these were the trusted source of information used for reporting deaths to the CQC and encompassed all residents. The number of unique residents present in the homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")` was estimated by multiplying the number of private residents (from the individual-level dataset) by the ratio of the total resident-days (from the weekly occupancy censuses) to the total private resident-days (from the individual data). This approach assumes that the mean length of stay of local authority residents is identical to that of private residents (supplementary material 1). Aggregate outbreak counts were also used to estimate incidence and cumulative incidence across time using a life table approach. Daily occupancy was inferred from weekly occupied bed censuses using linear interpolation. As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrements life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents (supplementary material 1). A Kaplan-Meier product limit estimator was applied to this life table to estimate the cumulative incidence of symptoms, confirmed infections, and Covid-19 related deaths every day between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. The incidence proportion of confirmed cases in England was contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020.[@cisbulletin2020; @cisdata2020]

>>>>

---
title: "Covid-19 infection and attributable mortality in residents of UK care homes: Cohort study using electronic records (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)
# override the finalfit cox PH modelling which defaults to Efron ties
# using Breslow ties instead
# coxphmulti <- function (.data, dependent, explanatory) {
#     requireNamespace("survival")
#     coxph(as.formula(paste0(dependent, "~", paste(explanatory, 
#         collapse = "+"))), data = .data, ties = "breslow")
# }
# assignInNamespace("coxphmulti",coxphmulti,ns="finalfit")

@@

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-14")
date_min_tallies = as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-14")
date_min_cox_mortality <- as.Date("2020-04-01")

@@

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Uninfected", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(covid_first_confirmed, na.rm = T), source="DAT")
) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(first_case_date = min(first_case_date)) 
  # tidyr::crossing(tibble(outbreak = c(0, 1)))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  ))  %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Uninfected", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_episodes <- dplyr::mutate(
  residents_episodes,
  covid_status_outbreak = as.factor(case_when(
    covid_status == "Uninfected" & outbreak == 0 ~ "0 Uninfected (other home)",
    covid_status == "Symptomatic (not confirmed)" &
      outbreak==0 ~ "1 Symptomatic not confirmed  (other home)",
    covid_status == "Uninfected" & outbreak == 1 ~ "2 Uninfected (outbreak home)",
    covid_status == "Symptomatic (not confirmed)" &
      outbreak==1 ~ "3 Symptomatic not confirmed  (outbreak home)",
    covid_status == "Asymptomatic confirmed" ~ "4 Confirmed asymptomatic",
    covid_status == "Symptomatic confirmed" ~ "5 Confirmed symptomatic"
  )))

residents_episodes$covid_status_outbreak_digit <- as.factor(
  as.numeric(residents_episodes$covid_status_outbreak)-1)
residents_episodes <- bind_cols(
  residents_episodes,
  as_tibble(model.matrix(~covid_status_outbreak_digit, residents_episodes)[ ,-1])
)

residents_episodes_short <- residents_episodes %>% 
  filter(exit_date > date_min_cox_mortality) %>% 
  mutate(date1 = if_else(
    entry_date > date_min_cox_mortality,
    date1,
    as.numeric( date_min_cox_mortality - date_min),
  ))

@@

incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Characteristics of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CFR_age_sex <- table_nums(
  name = "tab_CFR_age_sex",
  caption =  paste0("All-cause case-fatality rates and relative risks by age and sex among private residents (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  "; ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Risk factors for confirmed infection in private residents: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Risk factors for all-cause mortality in private residents of care homes with and without Covid-19 outbreaks: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_episodes_short$resident_id)),
  ", ", format(date_min_cox_mortality, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_attributable_deaths <- table_nums(
  name = "tab_attributable_deaths",
  caption =  paste0("Model-based estimates of attributable death in private residents of care homes with and without Covid-19 outbreaks (n=", 
  formatbm(dplyr::n_distinct(residents_episodes_short$resident_id)),
  ", ", format(date_min_cox_mortality, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

supp_epi_DAT <- table_nums(
  name = "supp_epi_DAT",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections amongst residents according to Datix incident reports (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

@@

tab_zero_cases <- table_nums(
  name = "tab_zero_cases",
  caption = paste0("Number and proportion of homes without confirmed cases by UK region (", format(date_min_tallies, "%e %b %Y"), "-", format(date_max_tallies, "%e %b %Y"), ")")
)

@@

fig_KM_CT_plots <- fig_nums(
  name = "fig_KM_CT_plots",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in (A) residents (n=",   formatbm(desc_residents$CT_total_residents),
  ") and (B) care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance counts (", format(date_min_tallies, "%e %b %Y"), "-", format(date_max_tallies, "%e %b %Y"), ")"))

@@

**Findings:** `r formatbm(desc_residents$CT_total_symptomatic)` residents developed Covid-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had confirmed infections. The incidence of confirmed infection in residents and staff respectively was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days.
`r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) had neither confirmed infections nor Covid-related deaths.

Among `r formatbm(desc_residents$DAT_total_residents)` private residents, there were `r formatbm(desc_residents$DAT_total_deaths)` all-cause deaths (`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%]), `r desc_residents$DAT_CFR$y` of which occurred in residents with confirmed infection, equivalent to an all-cause case-fatality rate of `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%]. `r tbrounding(((desc_residents$DAT_CFR_RR - 1)/desc_residents$DAT_CFR_RR)*100)`% [`r tbrounding(((desc_residents$DAT_CFR_RR_lower - 1)/desc_residents$DAT_CFR_RR_lower)*100)`%; `r tbrounding(((desc_residents$DAT_CFR_RR_upper - 1)/desc_residents$DAT_CFR_RR_upper)*100)`%] of these deaths were attributable to SARS-CoV-2. Symptomatic and asymptomatic infection were independent risk factors for mortality. All-cause mortality for uninfected residents was 93% higher in care homes with outbreaks compared to those without outbreaks (adjusted HR = 1·9 [1·7; 2·2]).

@@

Care home residents are particularly vulnerable to Covid-19 infection due to their advanced age and high prevalence of comorbidity,[@Zhou2020] and this is compounded by frequent exposure to infection through close contact with infected staff members, other residents and contaminated surfaces in the care home setting. In the UK, an estimated 420,000 residents lived in approximately 11,000 homes for the elderly.[@cqcdata2020]

At the peak of the pandemic the number of deaths in care home residents was three-fold higher than the equivalent period in in 2019.[@onsDR2020]  Data from the Office for National Statistics suggest that two-thirds of the 29,393 excess deaths recorded in care home residents between 28th December 2019 and 12th June 2020 (calculated relative to equivalent period in the previous year) involved Covid-19.[@onsDR2020] However, estimating the proportion of care home deaths that can be directly attributed to Covid-19 based on death certifications is challenging because many care home residents with suspected infection were not tested, particularly at the start of the pandemic. Understanding the ratio of direct to indirect deaths in care home residents will be critical to inform public health measures to mitigate the impact of the anticipated second wave of the pandemic, since lockdown measures that prevent infection may have a negative impact on residents by disrupting usual care.

Disentangling direct and indirect deaths from Covid-19 requires accurate information on the incidence and prevalence of infection among staff and residents linked to individual-level clinical outcomes. There is no syndromic surveillance for infection in care homes in England, and universal testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established in care homes in England until 11th May 2020, when the government launched the whole care home testing programme.[@dhscTesting2020] Prior to this, testing was only available for care home residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations with a maximum of five tests per care home. Consequently national estimates of incidence and prevalence based on test results obtained prior to 11th May will substantially underestimate the burden of infection in care home residents and staff.

Outbreak investigations provide estimates of the proportion of residents and staff who become infected during an outbreak and case-fatality rates,[@McMichael2020; @Dora2020] but they provide limited insight into the relationship between infection and mortality across the entire pandemic period, and cannot be generalised to care homes that have not reported any outbreaks. A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long-term care facilities. Mortality rates in residents with confirmed Covid-19 infection ranged from 0 to 10%, with case fatality rates between 0 and 34%, however this review did not include estimates of indirect mortality.

@@

## Aggregate outbreak surveillance counts (`r format(date_min_tallies, "%e %b %Y")`--`r format(date_max_tallies, "%e %b %Y")`)

@@

## Individual-level residents records (`r format(date_min, "%e %b %Y")`--`r format(date_max, "%e %b %Y")`)

@@

Incidence proportions and rates were calculated using the aggregate outbreak counts, as these were the trusted source of information used for reporting deaths to the CQC and encompassed all residents. The number of unique residents present in the homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")` was estimated by multiplying the number of private residents (from the individual-level dataset) by the ratio of the total resident-days (from the weekly occupancy censuses) to the total private resident-days (from the individual data). This approach assumes that the mean length of stay of local authority residents is identical to that of private residents (supplementary material 1). Aggregate outbreak counts were also used to estimate incidence and cumulative incidence across time using a life table approach. Daily occupancy was inferred from weekly occupied bed censuses using linear interpolation. As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrements life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents (supplementary material 1). A Kaplan-Meier product limit estimator was applied to this life table to estimate the cumulative incidence of symptoms, confirmed infections, and Covid-19 related deaths every day between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. The incidence proportion of confirmed cases in England was contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between `r format(england_results$date_min, "%e %B %Y")` and `r format(england_results$date_max, "%e %B %Y")`.[@cisbulletin2020; @cisdata2020]

Individual-level data (including dates of care home entry and exit) were available for private residents only. By assuming infection underreporting to affect residents conditionally at random, it was possible to use the data to study factors associated with the rate of infection and mortality. A Cox proportional hazards model was used to test the effect of both individual-level risk factors (sex, age, care type) and organisation-level risk factors (region, care home size, staff-to-bed ratio, occupant-to-bedroom ratio).
